### 2016-09-23 mixed-model.Rmd
preliminary look at mixed model & residuals

---
title: "R Notebook"
output: html_notebook
---

```{r, message = FALSE, warning = FALSE, results = "hide"}
devtools::load_all()
library(tidyverse)
library(ggplot2)
start_yr = 1982
end_yr = 2013
min_num_yrs = 25
last_train_year = 2003
d = get_richness_ts_env_data(start_yr, end_yr, min_num_yrs) %>% 
  filter(!is.na(observer_id))
train = d[d$year <= last_train_year, ]
test = d[d$year > last_train_year, ]
```

Partition the variance among site effects, observer effects, annual effects,
and residuals.

```{r}
library(lme4)
model = lmer(richness ~ (1|site_id) + (1|observer_id), data = train)
summary(model)
```

Predict forward:

```{r}
predictions = test %>%
  mutate(predicted = predict(model, test, allow.new.levels = TRUE)) %>% 
  mutate(diff = richness - predicted)

predictions %>% 
  group_by(year) %>% 
  summarize(mean(diff^2)) %>% 
  plot(ylab = "Mean Squared Errorr", type = "l", ylim = c(0, 50))

predictions %>% 
  filter(observer_id %in% train$observer_id) %>% 
  group_by(year) %>% 
  summarize(mean(diff^2)) %>% 
  lines(col = 2)
legend("topleft", 
       legend = c("all observations", "known observers only", "training observations"), 
       lty = c(1, 1, 2), col = c(1, 2, 1))

abline(h = var(resid(model)), lty = 2)
```

## Can we do better?

None of the predictor variables correlate with the residuals, suggesting that 
their explanatory effect is mainly *among* sites, rather than in the residuals
for individual sites.

```{r}
stopifnot(nrow(train) == length(resid(model)))
round(sort(cor(resid(model), train)[1, colnames(train) != "richness"]^2), 4)
```

There also isn't much in the way of consistent autocorrelation in the residuals
(correlation of 0.2 explains 4% of the residual variance [100 * 0.2^2], which 
isn't tremendously useful even if it is statistically significant---especially
since the residual variance is only a small portion of the total variance anyway).

```{r}
resid_matrix = data.frame(site_id = train$site_id, 
                          year = train$year, 
                          residuals = resid(model)) %>% 
  complete(site_id, year) %>% 
  arrange(site_id, year) %>% 
  magrittr::extract2("residuals") %>% 
  matrix(nrow = n_distinct(train$year))

autocorr_matrix = resid_matrix %>% 
  t() %>% 
  cor(use = "pairwise.complete.obs")

plot_upper = function(x, title){
  reshape2::melt(x * upper.tri(x)) %>% 
    ggplot(aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile() +
    scale_fill_gradient2() +
    ggtitle(title)
}

plot_upper(autocorr_matrix, "autocorrelation")

summary(lm(resid_matrix[4, ] ~ t(resid_matrix[1:3, ])))

```