### 2017-03-14 manuscript/important_references.Rmd

## Various factoids

#### On North American climate  
There has been an increase in mean temperature in North America since the 1980's (pg. 90, fig 2.7)   
Redmond, Kelly T., and John T. Abatzoglou. "Current climate and recent trends." Climate Change in North America. Springer International Publishing, 2014. 53-94.

Predictions for the next 10-20 years are uncertain (Meehl et al 2014) but models currently suggest a continued trend in warming, with potential for short term decreases (natural variability around increasing trend)(fig 3 & 4, Kumar et al 2014)  
Kumar, Devashish, Evan Kodra, and Auroop R. Ganguly. "Regional and seasonal intercomparison of CMIP3 and CMIP5 climate model ensembles for temperature and precipitation." Climate dynamics 43.9-10 (2014): 2491-2518.  
Meehl, Gerald A., et al. "Decadal climate prediction: an update from the trenches." Bulletin of the American Meteorological Society 95.2 (2014): 243-267.

-------------------------------
#### Biodiversity is an important thing to consider   
Cardinale, Bradley J., et al. "Biodiversity loss and its impact on humanity." Nature 486.7401 (2012): 59-67.

## Forcasting
#### People should do a better job at forecasting.  
Evans, M. R., Norris, K. J. & Benton, T. G. Predictive ecology: systems approaches. Philos. Trans. R. Soc. B Biol. Sci. 367, 163–169 (2012).  
Houlahan, J. E. The priority of prediction in ecological understanding. Oikos (2016). doi:10.1111/oik.03726  
Petchey, O. L. et al. The ecological forecast horizon, and examples of its uses and determinants. Ecol. Lett. 18, 597–611 (2015).

## SDM references

GBM's are one  of the best models for SDM's (Elith et al 2006) but JSDMs could potentially be better (Warton et al 2015)

Elith, J. et al. Novel methods improve prediction of species’ distributions from occurrence data. Ecography (Cop.). 29, 129–151 (2006).  
Warton, David I., et al. "So many variables: joint modeling in community ecology." Trends in ecology & evolution 30.12 (2015): 766-779.

## Bird ecology
Renfrew, R. B., Kim, D., Perlut, N., Smith, J., Fox, J. and Marra, P. P. (2013), Phenological matching across hemispheres in a long-distance migratory bird. Diversity Distrib., 19: 1008–1019. doi:10.1111/ddi.12080
Stephen Mayor, a postdoc with Rob Guralnick, also has unpublished work looking at migration/ndvi patterns for all north american birds.