### 2017-08-26 manuscript/draft.Rmd

---
title: "Best practices for forecasting changes in biodiversity: an example using breeding birds"
author:
- David J. Harris
- Shawn D. Taylor
- Ethan P. White
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse, quietly = TRUE)
library(kableExtra, quietly = TRUE)
```

`r settings = yaml::yaml.load_file("../settings.yaml")` 
`r timeframe = settings$timeframes$train_22` 
`r numbers = yaml::yaml.load_file("numbers.yaml")`

## Abstract

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. While each model could explain most of the variance
in richness, none of them consistently outperformed a baseline model that
predicted constant richness at each site. In particular, we found no evidence
that current methods (such as species distribution models) can successfully turn
spatial data into useful temporal predictions about biodiversity at decadale
time-scales. The best practices implemented in this study directly influence the
forecasts, the relative performance of different modeling approaches, and the
conclusions about the current state of biodiversity forecasting . To facilitate
the rapid improvement of biodiversity forecasts, we emphasize the value of
specific best practices in making forecasts and evaluating forecasting methods.

## Introduction

Forecasting the future state of ecological systems is increasingly important for
planning and management, and also for quantitatively evaluating how well
ecological models capture the key processes governing natural systems
[@clark2001; @houlahan2017; @dietze2017]. Forecasts regarding biodiversity are
especially important, due to biodiversity's central role in conservation
planning and its sensitivity to anthropogenic effects [@cardinale2012;
@diaz2015; @tilman2017]. High-profile studies forecasting large biodiversity
declines over the coming decades have played a large role in shaping ecologists'
priorities [as well as those of policymakers; e.g. @ipcc2014], but it is
inherently difficult to evaluate such long-term predictions before the projected
biodiversity declines have occurred.

Previous efforts to predict future patterns of species richness, and diversity
more generally, have focused primarily on building species distributions models
[SDMs; @thomas2004; @thuiller2011; @urban2015]. In general, these models
describe individual species' occurrence patterns as functions of the
environment. Given forecasts for environmental conditions, these models can 
predict where each species will occur in the future. These species-level 
predictions are then combined ("stacked") to generate forecasts for species 
richness [e.g. @calabrese2014]. Alternatively, models that directly relate spatial 
patterns of species richness to environment conditions have been developed and 
generally perform equivalently to stacked SDMs [@algar2009; @distler2015]. 
This approach is sometimes referred to as "macroecological" modeling, because it 
models the larger-scale pattern (richness) directly [@distler2015].

Despite the emerging interest in forecasting species richness and other aspects
of biodiversity [@jetz2007; @thuiller2011], little is known about how
effectively we can anticipate these dynamics. This is due in part to the long
time scales over which many ecological forecasts are applied [and the resulting
difficulty in assessing whether the predicted changes occurred; @dietze2016].
What we do know comes from a small number of hindcasting studies, where models
are built using data on species occurrence and richness from the past and
evaluated on their ability to predict contemporary patterns [e.g., @algar2009;
@distler2015]. These studies are a valuable first step, but lack several
components that are important for developing forecasting models with high
predictive accuracy, and for understanding how well different methods can
predict the future. These "best practices" for effective forecasting and
evaluation (Box 1) broadly involve: 1) expanding the use of data to include
biological and environmental time-series [@tredennick2016]; 2) accounting for
uncertainty in observations and processes, [@harris2015; @yu2010]; and 3)
conducting meaningful evaluations of the forecasts by hindcasting, archiving
short-term forecasts, and comparing forecasts to baselines to determine whether
the forecasts are more accurate than assuming the system is basically static
[@perretti2013].

In this paper, we attempt to forecast the species richness of breeding birds at
over 1,200 of sites located throughout North America, while following best
practices for ecological forecasting (Box 1). To do this, we combine 32 years of
time-series data on bird distributions from annual surveys with monthly
time-series of climate data and satellite-based remote-sensing. Datasets that
span a time scale of 30 years or more have only recently become available for
large-scale time-series based forecasting. A dataset of this size allows us to
model and assess changes a decade or more into the future in the presence of
shifts in environmental conditions on par with predicted climate change. We
compare traditional distribution modeling based approaches to spatial models of
species richness, time-series methods, and two simple baselines that predict
constant richness for each site, on average (Figure 1). All of our forecasting
models account for uncertainty and observation error, are evaluated across
different time lags using hindcasting, and are publicly archived to allow future
assessment. We discuss the implications of these practices for our understanding
of, and confidence in, the resulting forecasts, and how we can continue to build
on these approaches to improve ecological forecasting in the future.

![Example predictions from six forecasting models for a single site. Data from
1982 through 2003, connected by solid lines, were used for training the models;
the remaining points were used for evaluating the models' forecasts. In each
panel, point estimates for each year are shown with lines; the darker ribbon
indicates the 68% prediction interval (1 standard deviation of uncertainty), and
the lighter ribbon indicates the 95% prediction interval. **A.** Single-site
models were trained independently on each site's observed richness values. The
first two models ("average" and "naive") served as baselines. **B.** The
environmental models were trained to predict richness based on elevation,
climate, and NDVI; the environmental models' predictions change from year to
year as environmental conditions change.](../figures/model_predictions.png)

## Methods

We evaluated 6 types of forecasting models (Table 1) by dividing the 32 years
of data into 22 years of training data and 10 years of data for 
evaluating forecasts using hindcasting. We also made long term forecasts by using
the full data set for training and making forecasts through the year 2050. For
both time scales, we made forecasts using each model with and without
correcting for observer effects, as described below.

### Data

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@mcglinn2017]. The BBS data
was filtered to exclude all nocturnal, cepuscular, and aquatic species [since
these species are not well sampled by BBS methods; @hurlbert2005], as well as
unidentified species, and hybrids. All data from surveys that did not meet BBS
quality criteria were also excluded.

We used observed richness values from 1982 (the first year of complete
environmental data) to 2003 to train the models, and from 2004 to 2013 to test
their performance. We only used BBS routes from the continental United States
(i.e. routes where climate data was available @prism), and we restricted the
analysis to routes that were sampled during 70% of the years in the training
period (i.e., routes with at least 16 annual observations). The resulting
dataset included `r numbers$N_runs` annual surveys of `r numbers$N_sites` unique
sites, and included `r numbers$N_species` species.  Site-level richness varied
from `r numbers$richness_summary$min` to `r numbers$richness_summary$max` with
an average richness of `r numbers$richness_summary$mean` species.

**Past environmental data.** Environmental data included a combination of
elevation, bioclimatic variables and a remotely sensed vegetation index (the
normalized difference vegetation index; NDVI), all of which are known to
influence richness and distribution in the BBS data [@kent2014]. For each year
in the dataset, we used the 4 km resolution PRISM data [@prism] to calculate
eight bioclimatic variables identified as relevant to bird distributions
[@harris2015]: mean diurnal range, isothermality, max temperature of the warmest
month, mean temperature of the wettest quarter, mean temperature of the driest
quarter, precipitation seasonality, precipitation of the wettest quarter, and
precipitation of the warmest quarter. Satellite-derived NDVI, a primary
correlate of richness in BBS data [@hurlbert2002], was obtained from the NDIV3g
dataset with an 8 km resolution [@pinzon2014] and was available from 1981-2013.
Average summer (May, June, July) and winter (December, January, Feburary) NDVI
values were used as predictors. Elevation was from the SRTM 90m elevation
dataset [@jarvis2008] obtained using the R package raster [@raster].
Because BBS routes are 40-km transects rather than point counts, we used the
average value of each environmental variable within a 40 km radius of each BBS
route's starting point.

**Future environmental projections.** We made long term forecasts from 2014-2050
using the CMIP5 multi-model ensemble dataset as the source for climate
variables [@brekke2013]. Precipitation and temperature from 37 downscaled model
runs [@brekke2013, see Table S1] using the RCP6.0 scenario were averaged
together to create a single ensemble used to calculate the bioclimatic variables
for North America. For NDVI we used the per-site average values from 2000-2013
as a simple forecast. For observer effects (see below) each site was set to have
zero observer bias.

### Accounting for observer effects

Observer effects are inherent in large data sets collected by different
observers, and are known to occur in BBS [@sauer1994]. For each forecasting
approach, we trained two versions of the corresponding model: one with
corrections for differences among observers, and one without (Figure 2). We
estimated the observer effects (and associated uncertainty about those effects)
with a linear mixed model, with observer as a random effect, built in the Stan
probabilistic programming language [@carpenter2017]. Because observer and site are
strongly related (observers tend to repeatedly sample the same site), site was
also included as a random effect to ensure that inferred deviations were
actually observer-related (as opposed to being related to the sites that a given
observer happened to see). The resulting model partitions the variance in
observed richness values into site-level variance, observer-level variance, and
residual variance (e.g.  variation within a site from year to year). The
site-level estimates can also be used directly as the "average" baseline model
(see below). The estimated observer effects can be subtracted from the richness
values for a particular observer to provide an estimate of how many species
would have been found by a "typical" observer. To incorporate uncertainty in
these "corrected" richness values into the forecasting models we collected 500
Monte Carlo samples from the model's posterior distribution, and fit each of the
downstream models with each of the Monte Carlo samples. Each Monte Carlo sample
represented a different possible set of observer-level and site-level random
effect values across the full 32-year dataset.

![**A.** Model predictions for [[STATE]] route [[ROUTE NUMBER]] when all
observers are treated the same (black points). **B.** Model predictions for the
same route when accounting for systematic differences between observers
(represented by the points' colors). In this example most models are made more
robust to observer turnover. Note that the "naive" model is less sensitive to
observer turnover, and does not benefit as much from modeling
it.](../figures/observer_predictions.png)

### Models: site-level models

Three of the models used in this study were fit to each site separately, with no
environmental information (Table 1). These models were fit to each BBS route
twice: once using the residuals from the observer model, and once using the raw
richness values. When correcting for observer effects, we averaged across 500
models that were fit separately to the 500 Monte Carlo estimates of the observer
effects, to account for our uncertainty in the true values of those effects. All
of these models use a Gaussian error distribution (rather than a count
distribution) for reasons discussed below (see "Model evaluation").

\begin{singlespace}

```{r table, echo=FALSE}

table_caption = "Six forecasting models. Single-site models were trained
site-by-site, without environmental data. Environmental models were trained
using all sites together, without information regarding which transects occurred
at which site or during which year. Most of the models were trained to predict
richness directly. This mirrors the standard application of these
techniques. Separate random forest SDMs were fit for each species and used to
predict the probability of that species occurring at each site. The
species-level probabilities at a site were summed to predict richness. The
mistnet JSDM was trained to predict the full species composition at each site,
and the number of species in its predictions was used as an estimate of
richness."

table_data = read_csv("../figures/table.csv",col_types = "ccccc")

table_data %>%
  knitr::kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    caption = table_caption
  ) %>%
  add_header_above(c(" " = 2, "Predictors" = 3)) %>%
  group_rows("Single-site models", 1, 3) %>%
  group_rows("Environmental models", 4, 6) %>% kable_styling(position = "center")
```

\end{singlespace}

**Baseline models.** We used two simple baseline models as a basis for
comparison with the more complex models (Figure 2A). These baselines treated
site-level richness observations either as uncorrelated noise around a
site-level constant (the "average" model) or as an autoregressive model with a
single year of history [the "naive" model, @hyndman2014]. Predictions from the
"average" model are centered on the average richness observed during training,
and the confidence intervals are narrow and constant-width. The "naive" model,
in contrast, predicts that future observations will be similar to the final
observed value (e.g., in our hindcasts the value observed in `r
timeframe$last_train_year`), and the confidence intervals expand rapidly as the
predictions extend farther into the future. Both models' richness predictions
are centered on a constant value, so neither model can anticipate any trends in
richness or any responses to future environmental changes.

**Time series models.** We used Auto-ARIMA models [based on the `auto.arima`
function in @forecast1] to represent an array of different time-series modeling
approaches. These models can include an autoregressive component (as in the
"naive" model, but with the possibility of longer-term dependencies in the
underlying process), a moving average component (where the noise can have serial
autocorrelation) and an integration/differencing component (so that the analysis
could be performed on sequential differences of the raw data, accommodating more
complex patterns including trends). The `auto.arima` function chooses whether to
include each of these components (and how many terms to include for each one)
using AICc [@forecast1]. Since there is no seasonal component to the BBS
time-series, we did not include a season component in these models. Otherwise we
used the default settings for this function [@forecast1].

### Models: environmental models

In contrast to the single-site models, most attempts to predict species richness
focus on using correlative models based on environmental variables. We tested
three common variants of this approach: direct modeling of species richness;
stacking individual species distribution models; and joint species distribution
models (JSDMs). Following the standard approach, site-level random effects were
not included in these models as predictors, meaning that this approach
implicitly assumes that two sites with identical Bioclim, elevation, and NDVI
values should have identical richness distributions. As above, we included
observer effects and the associated uncertainty by running these models 500
times (once per MCMC sample).

**"Macroecological" model: richness GBM.** We used a boosted regression tree
model using the `gbm` package [Ridgeway *et al.* -@gbm] to directly model
species richness as a function of environmental variables. Boosted regression
trees are a form of tree-based modeling that work by fitting thousands of small
tree-structured models sequentially, with each tree optimized to reduce the
error of its predecessors. They are flexible models that are considered well
suited for prediction [@elith2008]. This model was optimized using a Gaussian
likelihood, with a maximum interaction depth of 5, shrinkage of 0.015, and up to
10,000 trees. The number of trees used for prediction was selected using the
"out of bag" estimator; this number averaged `r numbers$n_gbm[1]` for the
non-observer data and `r numbers$n_gbm[2]` for the observer-corrected data.

**Species Distribution Model: stacked random forests.** Species distribution
models (SDMs) predict individual species' occurrence probabilities using
environmental variables. Species-level models are used to predict richness by
summing the predicted probability of occupancy across all species at a site.
This avoids known problems with the use of thresholds for determining whether or
not a species will be present at a site [@calabrese2014; @pellissier2013].
Following @calabrese2014, we calculated the uncertainty in our richness estimate
by treating richness as a sum over independent Bernoulli random variables:
$\sigma^2_{richness} = \sum_i{p_i (1 - p_i)}$, where $i$ indexes species. By
itself, this approach is known to underestimate the true community-level
uncertainty because it ignores the uncertainty in the species-level probabilites
[@calabrese2014]. To mitigate this problem, we used an ensemble of 500 estimates
for each of the species-level probabilities instead of just one, propagating the
uncertainty forward. We obtained these estimates using random forests, a common
approach in the species distribution modeling literature. Random forests are
constructed by fitting hundreds of independent regression trees to
randomly-perturbed versions of the data [@cutler2007; @caruana2008]. When
correcting for observer effects, each of the 500 trees in our species-level
random forests used a different Monte Carlo estimate of the observer effects as
a predictor variable.

**Joint Species Distribution Model: mistnet.** Joint species distribution models
(JSDMs) are a new approach that makes predictions about the full composition of
a community instead of modeling each species independently as above
[@warton2015]. JSDMs remove the assumed independence among species and
explicitly account for the possibility that a site will be much more (or less)
suitable for birds in general (or particular groups of birds) than one would
expect based on the available environmental measurements alone. As a result,
JSDMs do a better job of representing uncertainty about richness than stacked
SDMs [@harris2015; @warton2015]. We used the `mistnet` package
[@harris2015] because it is the only JSDM that describes species' environmental
associations with nonlinear functions.

### Model evaluation

We defined model performance for all models in terms of continuous Gaussian
errors, instead of using discrete count distributions. Variance in species
richness within sites was lower than predicted by several common count models,
such as the Poisson or binomial (i.e. richness was underdispersed for individual
sites), so these count models would have had difficulty fitting the data [cf.
@calabrese2014]. The use of a continuous distribution is adequate here, since
richness had a relatively large mean (`r numbers$richness_summary$mean`) and all
models produce continuous richness estimates. When a model was run multiple
times for the purpose of correcting for observer effects, we used the mean of
those runs' point estimates as our final point estimate and we calculated the
uncertainty using the law of total variance (i.e. the average of the model runs'
variance, plus the variance in the point estimates).

We evaluated each model's forecasts using the data for each year between `r
timeframe$last_train_year + 1` and `r timeframe$end_yr`. We used three metrics
for evaluating performance: 1) root-mean-square error (RMSE) to determine how
far, on average, the models' predictions were from the observed value; 2) the
95% prediction interval coverage to determine how well the models predicted the
range of possible outcomes; and 3) deviance (i.e. negative 2 times the Gaussian
log-likelihood) as an integrative measure of fit incorporating good point
estimates, precision, and coverage. In addition to evaluating forecast
performance in general, we evaluated how performance changed as the time horizon
of forecasting increased by plotting performance metrics against year. Finally,
we decomposed each model's squared error into two components: the squared error
associated with site-level means and the squared error associated with annual
fluctuations in richness within a site. This decomposition describes the extent
to which each model's error depends on consistent differences among sites versus
changes in site-level richness from year to year.

All analyses were conducted using R [@R]. Primary R packages used in the
analysis included dplyr [@dplyr], tidyr [@tidyr], gimms [@gimms], sp [@sp1;
@sp2], raster [@raster], prism [@prism], rdataretriever [@rdataretriever],
forecast [@forecast1; @forecast2], git2r [@git2r], ggplot [@ggplot2], mistnet
[@harris2015], viridis [@viridis], rstan [@rstan], yaml [@yaml], purrr [@purrr],
gbm [Ridgeway *et al.* -@gbm], randomForest [@randomForest]. Code to fully
reproduce this analysis is available on GitHub
(https://github.com/weecology/bbs-forecasting) and archived on Zenodo
(https://doi.org/10.5281/zenodo.839581).

## Results

The site-observer mixed model found that 70% of the variance in richness in the
training set could be explained by differences among sites, and 21% could be
explained by differences among observers. The remaining 9% represents residual
variation, where a given observer might report a different number of species in
different years. In the training set, the residuals had a standard deviation of
about `r numbers$resid_sd` species. After correcting for observer differences,
there was little temporal autocorrelation in these residuals (i.e. the residuals
in one year explain `r round(100 *numbers$residual_autocor^2, 1)`% of the
variance in the residuals of the following year), suggesting that richness was
approximately stationary between 1982 and 2003.

When comparing forecasts for richness across sites all methods performed well
(Figure 3; all $R^2 > 0.5$). However SDMs (both stacked and joint) and the
macroecological model all failed to successfully forecast the highest-richness
sites, resulting in a notable clustering of predicted values near ~60 species
and the poorest model performance ($R^2$=`r numbers$env_R2s`, versus $R^2$=`r
numbers$ts_R2s` for the within-site methods).

![Performance of six forecasting models for predicting species richness one year
(2004) and ten years into the future (2013). Plots show observed vs. predicted
values for species richness. Models were trained with data from 1982-2003.  In
general, the single-site models (**A**) outperformed the environmental models
(**B**). The accuracy of the predictions generally declined as the timescale of
the forecast was extended from 2004 to 2013.](../figures/scatter.png)

While all models generally performed well in absolute terms (Figure 3), none
consistently outperformed the "average" baseline (Figure 4). The auto-ARIMA was
generally the best-performing non-baseline model, but in many cases (67% of the
time), the auto.arima procedure selected a model with only an intercept term
(i.e. no autoregressive terms, no drift, and no moving average terms), making it
similar to the "average" model. All five alternatives to the "average" model
achieved lower error on some of the sites in some years, but each one had a
higher mean absolute error and higher mean deviance (Figure 4).

![Difference between the forecast error of models and the error of the average
baseline using both absolute error (**A.**) and deviance (**B.**). Differences
are taken for each site and testing year so that errors for the same forecast
are directly compared. The error of the average baseline is by definition zero
and is indicated by the horizontal gray line.  None of the five models provided
a consistent improvement over the average baseline. The absolute error of the
models was generally similar or larger than that of the "average" model, with
large outliers in both directions.  The deviance of the models was also
generally higher than the "average" baseline.](../figures/model_violins.png)

![Change in performance of the six forecasting models with the time scale of the
forecast (1-10 years into the future). **A.** Root mean square error (rmse; the
error in the point estimates) shows the three environmental models tending to
show the largest errors at all time scales and the models getting worse as they
forecast further into the future at approximately the same rate. **B.** Deviance
(lack of fit of the entire predictive distribution) shows the stacked species
distribution models with much higher error than other models and shows that the
"naive" model's deviance grows relatively quickly. **C.** Coverage of a model's
95% confidence intervals (how often the observed values fall inside the
predicted range; the black line indicates ideal performance) shows that the
"naive" model's predictive distribution is too wide (capturing almost all of the
data) and the stacked SDM's predictive distribution is too narrow (missing
almost a third of the observed richness values by
2014).](../figures/performance_time.png)

Most models produced confidence intervals that were too narrow, indicating
overconfident predictions (Figure 5C). The random forest-based SDM stack was the
most overconfident model, with only `r numbers$rf_coverage_pct`% of observations
falling inside its 95% confidence intervals. This stacked SDM's narrow
predictive distribution caused it to have notably higher deviance (Figure 5B)
than the next-worst model, even though its point estimates were not unusually
bad in terms of RMSE (5A). As discussed elsewhere [@harris2015], this
overconfidence is a product of the assumption in stacked SDMs that errors in the
species-level predictions are independent. The GBM-based "macroecological" model
and the mistnet JSDM had the best calibrated uncertainty estimates (Figure
5B)and therefore their relative performance was higher in terms of deviance than
in terms of RMSE. The "naive" model was the only model whose confidence
intervals were too wide (Figure 5C), which can be attributed to the rapid rate
at which these intervals expand (Figure 1).

Partitioning each model's squared error into consistent errors across sites and
errors that fluctuated year-to-year shows that the majority of the error comes
from errors in the site-level mean (Figure 6). The "average" model, which was
based on the site-level mean, had the lowest error in this regard. In contrast,
the three environmental models showed the largest biases at the site level; this
makes sense, given that they could not explicitly distinguish among sites with
similar climate, NDVI, and elevation.

![Partitioning of the squared error for each model into site and year
components. The site-level mean component shows consistent over or under
estimates of richness at a site across years. The annual fluctuation compoonent
shows errors in predicting fluctuations in a site's richness over time. Both
components of the mean squared error were lower for the single-site models than
for the environmental models.](../figures/barcharts.png)

Accounting for differences among observers generally improved measures of model
fit (Figure 7). Improvements primarily resulted from a small number of forecasts
where observer turnover caused a large shift in the reported richness
values. The naive baseline was less sensitive to these shifts, because it
largely ignored the richness values reported by observers that had retired by
the end of the training period (Figure 1). The average model, which gave equal
weight to observations from the whole training period, showed a larger decline
in performance when not accounting for observer effects -- especially in terms
of coverage. The performance of the mistnet JSDM was notable here, because its
prediction intervals retained good coverage even when not correcting for
observer differences, which we attribute to the JSDM's ability to model this
variation with its latent variables.

![Accounting for differences between observers typically has a small impact on
accuracy (especially for the "naive" model), but occasonally changes a
prediction's accuracy by more than 10 species. As a result the inclusion of
observer effects results in average improvements in overall model
fit.](../figures/observers.png)

## Discussion

Forecasting is an emerging imperative in ecology; as such, the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts. We have laid out ideas for a number of these practices (Box 1) and
attempted to implement them in a single study that builds and evaluates
forecasts of biodiversity in the form of species richness. The results of this
effort are both promising and humbling. When comparing forecasts across sites,
many different approaches to forecasting produce reasonable forecasts. If a site
is predicted to have a high number of species in the future, relative to other
sites, it generally does. However (after controlling for observer differences),
none of the methods we evaluated could reliably tell us whether site-level
richness would increase or decrease over time (Figure 6), which is generally the
stated purpose of these forecasts. As a result, our baseline models, which did
not attempt to anticipate changes in richness over time, generally provided the
best forecasts for future biodiversity.

These results raise a number of important issues. First, we should be skeptical
of predictions about future biodiversity from models that have not been
validated in terms of repeated observations at the same site. Since site-level
richness is relatively stable, model evaluations that show high
accuracy across spatial gradients may not indicate much about a model's ability
to predict changes over time. This issue has also recently been highlighted as a
potential downside of using SDMs to predict the future locations of individual
species [@rapacciuolo2012; @oedekoven2017]. This result is particularly sobering
because SDMs form the main foundation for estimates of the predicted loss of
biodiversity to climate change [@thomas2004; @thuiller2011; @urban2015].

Second, our results highlight the importance of comparing multiple modeling
approaches when conducting ecological forecasts, and in particular, the value of
comparing results to simple baselines to avoid over-interpreting the information
present in these forecasts (Box 1). Disciplines that have more mature
forecasting cultures often do this by reporting "forecast skill", i.e., the
improvement in the forecast relative to a simple baseline [@jolliffe2003]. We
recommend following the advice of @ye2015 and adopting this approach in future
ecological forecasting research. Our results also highlight the need for using
hindcasting and time-series data for assessing the effectiveness of forecasting
methods. This practice allows comparisons to baseline models and also
facilitates the assessment of predicted dynamics within sites or regions.

One of the main alternatives to stacked SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts. Two
studies compared the SDM and direct richness modeling approaches, and reported
that the methods yielded equivalent results for forecasting diversity
[@algar2009; @distler2015]. While our results generally support the idea of
rough equivalence between stacked SDMs and direct richness modeling for point
estimates, they also show that stacked SDMs dramatically understimate the
uncertainty. For this reason, direct modeling of richness is a better overall
approach to forecasting richness. A similar result is seen when comparing joint
species distribution models (JSDMs) to stacked single species distribution
models. For point estimates, the joint distribution models are roughly
equivalent to stacked SDMs, but the JSDMs provide much better estimates of
uncertainty [@harris2015; @warton2015]. In fact, JSDMs and direct richness modeling
provide some of the best estimates of uncertainty across all modeling
approaches. This highlights the importance of evaluating models based on their
uncertainty as well as their mean estimates (Box 1). Not doing so is
particularly problematic in this case, because the summed species distribution
models are overly confident, which means that these models could mislead
ecologists into believing that richness would be restricted to a much narrower
range than would actually occur.

[[Note: this paragraph may no longer be accurate!]] For the error metric that
evaluates both models' point estimates and their uncertainty simultaneously, the
JSDM and the nonlinear richness model converge towards the time-series baselines
at the longest time-scales assessed. This convergence at longer time scales
highlights the importance of considering how forecasting method performance
changes with the distance into the future being predicted (Box 1). The value of
this approach, and associated forecast horizons, has been raised for knowing how
far into the future a model can be effectively used for forecasting
[@petchey2015], but our results suggest a broader value to this approach for
considering the potential importance of different models and processes for
making forecasts at different scales. Our results show that for forecasts 1-3
years in the future, baselines and time-series based approaches outperform all
other methods using integrative metrics. At these time scales, environmental
changes are relatively small and changes in the biota may lag behind changes in
the environment. However, as the timescale of the forecast increases to a decade
the JSDMs and environmental richness models get closer to the accuracy of the
time-series methods and even surpass some of those methods. As the forecast
distance increases the amount of environmental change is expected to increase
and the system will have more time to respond, potentially leading to shifts
towards improved relative performance of the models incorporating environmental
and ecological information. If this shift with forecast distance in the relative
importance of different processes and models continues, this could lead to JSDMs
and environmental richness models outperforming time-series based approaches at
sufficiently long time scales. These results suggest that assessment of how
forecast performance changes with time lag is not only important for determining
how far in the future to forecast, but also for determining which models and
processes are most relevant for making forecasts at different time scales. For
example ensemble forecasts could be created which incorporate these time lags
into the weights for individual models.

It is also possible that models that include species' relationships to their
environments or direct environmental constraints on richness will continue to
produce forecasts that are worse than simply assuming the systems are static.
This would be expected to occur if richness in these systems is not changing
over the relevant time scales (e.g. 10-30 years in this study), which would make
simpler models with no directional change more appropriate. Recent suggestions
that local scale richness in some systems is not changing directionally at
multi-decadal scales supports this possibility [@brown2001; @ernest2001;
@dornelas2014; @vellend2013]. This lack of change in richness may be expected
even in the presence of substantial changes in environmental conditions and
species composition at a site due to the replacement of species from the
regional pool [@brown2001; @ernest2001]. On average, the Breeding Bird Survey
sites used in this study show little change in richness [site-level SD of `r
numbers$resid_sd` species, after controlling for differences among observers;
see also @lasorte2005]. The absence of rapid change in this dataset is
beneficial for the absolute accuracy of forecasts across different sites: when a
past year's richness is already known, it is easy to estimate a future richness.
The site-level stability of the BBS data also explains why stacked SDMs perform
relatively well at predicting future richness, despite failing to capture
changes in richness over time. However, this stability also makes it difficult
to improve forecasts relative to simple baselines, since those baselines are
already close to representing what is actually occurring in the system. These
results suggest that single-site models should be actively considered for
forecasts of richness and other stable aspects of biodiversity. Our results also
suggest that future efforts to understand and forecast biodiversity should also
focus on species composition, since lower-level processes are expected to be
more dynamic [@ernest2001; @dornelas2014] and contain more useful information
[@harris2015].

**Observer effects.** In addition to consideration of the different process
models used for forecasting it is important to consider the observation models.
When working with any ecological dataset, there are imperfections in the
sampling process that have the potential to influence results. With large scale
surveys and citizen science datasets, such as the Breeding Bird Survey, these
issues are potentially magnified by the large number of different observers and
by major differences in the habitats and species being surveyed [@sauer1994].
Accounting for differences in observers reduced the average error in our point
estimates and also improved the coverage of the confidence intervals. In
addition, controlling for observer effects resulted in changes in which models
performed best, most notably reducing the relative performance of the naive
model's point estimates. This suggests that, prior to accounting for observer
effects, the naive model performed well because it was capable of accommodating
rapid shifts in estimated richness introduced by changes in the observer. These
kinds of rapid changes were difficult for the other single-site models to
accommodate, and so the average performance of the ARIMA and average models
improved once this source of observation error was addressed. This demonstrates
that modeling observation error can be important for properly estimating and
reducing uncertainty in forecasts and can also lead to changes in the best
methods for forecasting. We did not address differences in detection probability
across species and sites [@boulinier1998] since there is no clear way to address
this issue without making strong assumptions about the data (i.e., assuming
there is no biological variation in stops along a route), but this would be a
valuable addition to future forecasting models.

**Environmental uncertainty.** Future biodiversity forecasting efforts will also
need to begin to address the additional uncertainty introduced by the error
inherent in the process of forecasting the environmental conditions that
ecologists use as predictor variables. In this, and other hindcasting studies,
the environmental conditions for the "future" are known because the data has
already been observed. However, in real forecasts the environmental conditions
themselves have to be predicted, and environmental forecasts will also have
uncertainty and bias. Ultimately, ecological forecasts that use environmental
data will therefore be more uncertain than our current hindcasting efforts.
Thus, it is important to correctly incorporate this uncertainty in the predictor
variables into forecasting models [@clark2001; @dietze2017], as we did with the
uncertainty in observer-level effects. Difficulty in forecasting future
environmental conditions at small scales will present continued challenges for
models incorporating environmental variables, and this may result in a continued
advantage for simple single-site approaches.

**Multi-timescale forecasts.** Another challenge for ecological forecasting is
the timescales of typical forecasts. We evaluated forecasts using yearly
timesteps up to a decade into the future. In contrast, many ecological forecasts
are temporally aggregated to 5-30 year timesteps and projected up to a century
into the future. These are commonly beyond the career or even lifespan of the
researchers, which makes it impossible to directly validate the models'
predictions. Currently, assessments at these long time-scales can only be made
with a small number of opportune datasets. For example, hindcasts have been used
to assess models of species richness [@algar2009; @distler2015] and distribution
[@rapacciuolo2012; @moran-ordonez2017; @araujo2005; @eskildsen2013] using data
aggregated across similar time scales. More evaluation
studies like these are needed but there is a paucity of adequate long term data
available. Short term forecasts will allow for relatively rapid validation of
models by researchers without needing to wait for enough data to accumulate
[@tredennick2016; @dietze2016]. This does not mean that they should replace
large scale long-term forecasts. As discussed above, drivers of species richness
are likely to differ at different temporal scales [@rosenzweig1995; @white2004;
@white2007]. Thus, short-term forecasts should be expected to add value to
scientists and policymakers, and inform (not replace) long-term forecasts.
Hopefully, by evaluating trends in the shorter term forecasts as done here it
will be possible to also use the information from short-term forecasting to
improve longer-term estimates.

**Conclusions.** The science of forecasting biodiversity, and ecology more
broadly, remains in its infancy and it is important to consider the general
inability of forecasting methods to improve on simple baselines in that context.
When weather forecasting first started the forecasts were likewise worse than
simple baselines [@mcgill2012]. One practice that helped weather forecasts to
improve was publicly making a large number of predictions, which allowed
different forecasting approaches to be regularly assessed and refined
[@mcgill2012; @silver2012]. This suggests that it is important for ecologists to
start regularly making and evaluating ecological forecasts on real data, even if
they perform poorly, and to make these forecasts publicly available for
assessment. These forecasts should include both short-term predictions, which
can be assessed quickly, and mid- to long-term forecasts, which can help
ecologists to assess long time-scale processes and determine how far into the
future we can successfully forecast [@tredennick2016; @dietze2017]. Forecasts
based on the models in this paper (from now until 2050) are openly archived on
Zenodo, so that we and others can assess how well they perform. We also plan to
evaluate these forecasts and report the results as each new year of BBS data
becomes available at the same URL [@zenodo]. Weather forecasting has continually
improved throughout its history [@bauer2015], in part due to making and
evaluating public forecasts [@mcgill2012], and we hope this practice will help
ecology do the same.

Making successful ecological forecasts will be challenging. Ecological systems
are complex, fundamental theory is less refined than simpler physical/chemical
systems, and we currently lack the large amounts of data that produce effective
forecasts through machine learning. Despite this, we believe that progress can be
made if we build an active forecasting culture in ecology that builds and
assesses forecasts in ways that will allow us to improve the effectiveness of
ecological forecasts most rapidly (Box 1). This includes expanding the scope of
the ecological and environmental data we work with, paying attention to
uncertainty in both model building and forecast evaluation, and rigorously
assessing forecasts using a combination of hindcasting, archived forecasts, and
comparisons to simple baselines.

## Acknowledgments

This research was supported by the Gordon and Betty Moore Foundation's
Data-Driven Discovery Initiative through Grant GBMF4563 to E.P. White. We thank
the developers and providers of the data and software that made this research
possible including: the PRISM Climate Group at Oregon State University, the
staff at USGS and volunteer citizen scientists associated with the North
American Breeding Bird Survey, NASA, the World Climate Research Programme's
Working Group on Coupled Modelling and its working groups, the U.S. Department
of Energy's Program for Climate Model Diagnosis and Intercomparison, and the
Global Organization for Earth System Science Portals. A. C. Perry provided
valuable comments that improved the clarity of this manuscript.

## Box 1: Ten simple rules for making and evaluating ecological forecasts

### 1. Compare multiple modeling approaches

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process-based [e.g.
@kearney2009] and data-driven models, as well as comparing the accuracy of
forecasts to simple baselines to determine if the modeled forecasts are more
accurate than the naive assumption that the world is static [@ye2015].

### 2. Use time-series data when possible

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for-time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process and error structure perspective [@tredennick2016].

### 3. Pay attention to uncertainty

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in highly uncertain outcomes leading to poor decision making
and erosion of confidence in ecological forecasts. Models should explicitly
include sources of uncertainty and propagate them through the forecast where
possible [@clark2001; @dietze2017]. Evaluations of forecasts should assess the
accuracy of models' estimated uncertainties as well as their point estimates
[@dietze2017].

### 4. Use predictors related to the question

Many ecological forecasts use data that is readily available and easy to work
with. While ease of use is a reasonable consideration it is also important to
include predictor variables that are expected to relate to the ecological
quantity being forecast. Time-series of predictors, instead of long-term
averages, are also preferable to match the ecologial data (see #2).
Investing time in identifying and acquiring better predictor
variables may have at least as many benefits as using more sophisticated
modeling techniques [@kent2014].

### 5. Address unknown or unmeasured predictors

Ecological systems are complex and many biotic and abiotic aspects of the
environment are not regularly measured. As a result, some sites may deviate in
consistent ways from model predictions. Unknown or unmeasured predictors can be
incorporated in models using site-level random effects (potentially spatially
autocorrelated) or by using latent variables that can identify unmeasured
gradients [@harris2015].

### 6. Assess how forecast accuracy changes with time-lag

In general, the accuracy of forecasts decreases with the length of time into the
future being forecast [@petchey2015]. This decay in accuracy (and the potential
for different rates of decay to result in different relative model performance
at different lead times) should be considered when evaluating forecasts
and comparing models.

### 7. Include an observation model

Ecological observations are influenced by both the underlying biological
processes (e.g. resource limitation) and how the system is sampled. When
possible, forecasts should model the factors influencing the observation of the
data. [@yu2010; @hutchinson2011; @schurr2012]

### 8. Validate using hindcasting

Evalutating a model's predictive performance across time is critical.
Hindcasting uses a temporal out-of-sample validation approach to mimic how well
a model would have performed had it been run in the past. For example, using
occurance data from the early 20th century to model distributions which are
validated with late 20th century occurances. Dense time series, such as yearly
observations, are desirable to also evalulate the forecast horizon (see #6), but
this is not a strict requirement.

### 9. Publicly archive forecasts

Forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012]. Enough information
should be provided in the archive to allow ecologists to make an unambiguous
assessment of each forecast's performance.

### 10. Make both short-term and long-term predictions

In cases where long-term predictions are the primary goal, short-term
predictions should also be made to accommodate the time-scales of planning and
management decisions and to allow the accuracy of the forecasts to be quickly
evaluated [@tredennick2016].

## References