### 2017-12-07 manuscript/observer-equations.Rmd

---
output: pdf_document
fontsize: 10pt
geometry: left=2.5cm, right=2.5cm, top=1.5cm, bottom=2cm, headheight=0pt, letterpaper
---

\begin{center}

\textbf{Supplement to "Forecasting biodiversity in breeding birds using best practices"}

by David J. Harris, Shawn D. Taylor, and Ethan P. White

\end{center}

\raggedright

\vspace{3mm}

**Observer model description**

Richness at site $i$, as recorded by observer $j$ is estimated using a linear mixed model.  Thus, the response variable was modeled as $y_{ij} \sim \mathcal{N}(\mu_{ij}, \sigma^\mathrm{residual}).$

Here, $\mu_{ij}$ is defined as an intercept term, plus a site-level effect, plus an observer-level effect: $\mu_{ij} = \alpha + \alpha^\mathrm{site}_i + \alpha^\mathrm{observer}_j.$

These two effects are each drawn from zero-mean Gaussians: $\alpha^\mathrm{site} \sim \mathcal{N}(0, \sigma^\mathrm{site})$ and $\alpha^\mathrm{observer} \sim \mathcal{N}(0, \sigma^\mathrm{observer}).$

Prior distributions on $\alpha$, $\sigma^\mathrm{residual}$, $\sigma^\mathrm{site}$, and $\sigma^\mathrm{observer}$ are included below.

\vspace{3mm}

**Stan code**

```{text, code=readLines("../mixed.stan")}
```